### 2020-05-15 writeup/proof_transfer_ex.tex
sec 3.1
A.2
update

\subsection{Proof of Proposition \ref{prop_dist_transition}}

@@

>>>>

\subsection{Proof of Proposition \ref{prop_dist_transition}}\label{app_proof_31}

@@

\subsection{Proof of Proposition \ref{prop_data_size}}

\subsection{Proof of ???}
Note that for the case of $k$ tasks with the same covariates, since there is no covariate shift and the data ratio is always equal to one, the main factor is model distance.

\paragraph{A precise bound when there is no model shift.}
As Proposition \ref{prop_monotone} shows, if $\beta_s$ and $\beta_t$ are equal, then adding the source task dataset always helps learn the target task.
The goal of this section is to understand how covariate shift affects the rate of transfer. \todo{add conceptual msg}

%The key quantity is to look at:
%The estimator using the source and target together from minimizing \eqref{eq_mtl_basic} is
%\[ \hat{\beta}_{s,t} = (X_1^{\top} X_1 + X_2^{\top} X_2)^{-1} (X_1^{\top}Y_1 + X_2^{\top}Y_2)\]
%The estimation error of $\hat{\beta}_{s,t}$ is
%\begin{align}\label{eq_two_task}
%  \err(\hat{\beta}_{s,t}) = \sigma^2 \cdot \tr[(X_1^{\top}X_1 + X_2^{\top} X_2)^{-1}].
%\end{align}
%The estimation error using the target alone is
%\begin{align}\label{eq_target_task}
%	\err(\hat{\beta}_t) = \sigma^2 \cdot \tr[(X_2^{\top} X_2)^{-1}].
%\end{align}
%The improvement of estimation error from adding the source task is then given by
%$\err(\hat{\beta}_t) - \err(\hat{\beta}_{s,t})$.
%For the test error on the target task, the improvement from adding the source task is
%\[ \te(\hat{\beta}_t) - \te(\hat{\beta}_{s,t}) = \sigma^2\cdot\bigtr{\bigbrace{(X_2^{\top}X_2)^{-1} - (X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}\cdot\Sigma_2}. \]

%We calculate the amount of improvement by comparing equation \eqref{eq_two_task} to equation \eqref{eq_target_task}.
A simple observation here is that when $\beta_s = \beta_t$, the optimal $\hat{w}$ for minimizing equation \eqref{eq_te_mtl} is equal to $1$.
Based on this observation, we can get a more precise result than Theorem \ref{thm_model_shift} on the improvement of adding the source task data that only depends on the covariance matrices $\Sigma_1, \Sigma_2$ and the number of data points $n_1, n_2$.

\begin{proposition}[Transfer rate without model shift]\label{thm_cov_shift}
Suppose $\beta_s = \beta_t$ and $\|\beta_t\|_2^2\sim p\sigma^2$ (i.e. the $l^2$-norm of the vector $\beta_t$ is of the same order as that of the error vector). Assume that the condition numbers of $\Sigma_1$, $\Sigma_2$ and $M:=\Sigma_1^{1/2}\Sigma_2^{-1/2}$ are all bounded by a constant $C>0$. Then we have that the optimal ratio for $W_1/W_2$ in equation \eqref{eq_te_mtl} satisfies
	$$1\le \hat{w} \le 1+\OO(p^{-1}).$$%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
	%where $M:=\Sigma_1^{1/2}\Sigma_2^{-1/2}$.
Moreover, we have
	\begin{align}\label{tehatw1}
		%\err(\hat{\beta}^{\TL}_{s,t}) &= \sigma^2 \cdot \bigtr{\frac 1 {(n_1 + n_2)a_1\Sigma_1 + (n_1 + n_2)a_2\Sigma_2}} \\
		\te(\hat{\beta}^{\TL}_{t}) &= \sigma^2 \cdot \bigtr{\bigbrace{(n_1 + n_2)a_1 M^\top M  + (n_1 + n_2)a_2\id}^{-1}} \cdot \left(1+ \bigo{p^{-1}}\right),
	\end{align}
where $a_1, a_2$ are the solutions to equations \eqref{eq_a2}. %\cor $w_0$ is close to 1 if the signal strength $\beta_t$ is much larger than the noise strength \nc
\end{proposition}

\begin{proof}
We abbreviate $\val(w_2\hat{B}(w)):=\val(w)$. Note that $\val(w)\le \val(-w)$ for $w\ge 0$. Hence we have $\hat w\ge 0$. Moreover, we notice that $\val (w) < \val (1)$ for all $0\le w < 1$. Thus we have $\hat w\ge 1$. It suffices to consider the case with $w> 1$. Under the assumption on $\beta_s$ and $\beta_t$, we can write
\begin{align}
	\val(w) =&~  \left( 1-\frac1w\right)^2 \left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2 \nonumber \\
			&~ + \frac{\sigma^2}{w^2} \cdot \bigtr{( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} }. \nonumber
\end{align}
Since
\begin{align*}
&\left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2 \\
&= \tr \left[ ( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-2}M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t\beta_t^\top \Sigma_1^{1/2}Z_1^{\top} Z_1  M\right]
\end{align*}
is increasing with respect to $w$, then the derivative of $\val(w)$ can be bounded from below as
\begin{align*}
\val'(w) \ge &~ 2\frac{w-1}{w^3} \left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2   \\
			&~ - 2 \frac{\sigma^2}{w^3} \cdot \bigtr{(M^\top Z_1^{\top}Z_1 M +w^{-2}  Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top}Z_1 M (M^\top Z_1^{\top}Z_1 M + w^{-2} Z_2^{\top}Z_2)^{-1}} \\
\ge &~ 2\frac{w-1}{w^3} \left\|( M^\top Z_1^{\top}Z_1 M +  Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2   - 2 \frac{\sigma^2}{w^3} \cdot \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}}.
			%\\ =& ~ 2\frac{d^2}{w^3} \tr\left[( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} \left[\left( w-1\right)\left(Z_1 \Sigma_1 Z_1^{\top}\right) - \frac{\sigma^2}{d^2}\id \right] Z_1 M ( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1}\right]  .
\end{align*}
Hence $\val'(w)\ge 0$ if $w>1+ \e_0$, where
$$\e_0:= \frac{\sigma^2  \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}} }{\left\|( M^\top Z_1^{\top}Z_1 M + Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2}.$$
In other words, $\val(w)$ is strictly increasing function on $[1+\e_0,\infty]$. Thus we get that $\hat w$ satisfies
\be\label{hatw1}1\le w \le 1+\e_0.\ee Using \eqref{eq_isometric}, we get that
$$\e_0=\OO(\sigma^2/\|\beta_t\|_2^2)=\OO(p^{-1}).$$

Finally, plugging \eqref{hatw1} into the expression $\te(\hat{\beta}^{\TL}_{t}) $, we obtain \eqref{tehatw1}.
%$$1\le \hat{w} \le w_0:=1 +\frac{\sigma^2  \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}} }{\left\|( M^\top Z_1^{\top}Z_1 M + Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2},$$
\end{proof}

As a remark, we see that Proposition \ref{prop_monotone} follows from Theorem \ref{thm_cov_shift}.
The amount of reduction on test error for the target task is given as
	\begin{align*}
%		\err(\hat{\beta}_t) - \err(\hat{\beta}_{s,t})
%		&= \sigma^2 p \cdot \bigtr{\frac 1 {(n_2 - p) \Sigma_2} - \frac 1 {(n_1 + n_2)a_1 \Sigma_1 + (n_1 + n_2)a_2 \Sigma_2}}, \\
		\te(\hat{\beta}_t) - \te(\hat{\beta}_{s,t})
		&= \sigma^2 \cdot \bigbrace{\frac p {n_2 - p} -  \bigtr{\bigbrace{(n_1 + n_2)a_1\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2} + (n_1 + n_2)a_2\id}^{-1}}}.
	\end{align*}
Because
\begin{align*}
	\te(\hat{\beta}_{s,t}) \le \te(\hat{\beta}_t)
	\Leftarrow~ & (n_2 - p)\Sigma_2 \preceq (n_1 + n_2) a_1 \Sigma_1 + (n_1 + n_2)a_2 \Sigma_2 \\
	\Leftrightarrow~ & \zeroMatrix \preceq (n_1 + n_2) a_1 \Sigma_1 + (n_1 - (n_1 + n_2)\cdot a_1) \Sigma_2,
\end{align*}
which is true since $a_1 \le n_1 / (n_1 + n_2)$ by equation \eqref{eq_a2}.
The proof for $\te(\hat{\beta}_{s,t}) \le \te(\hat{\beta}_t)$ follows by multiplying $\Sigma_2^{-1/2}$ on both sides of the inequalities above.

\medskip

### 2020-05-15 writeup/transfer.tex
edits
update

\subsection{Task Model Similarities versus Noise}

@@

@@

%\subsubsection{Gains over Single-Task Training}

Note that for the case of $k$ tasks with the same covariates, since there is no covariate shift and the data ratio is always equal to one, the main factor is model distance.

\paragraph{A precise bound when there is no model shift.}
As Proposition \ref{prop_monotone} shows, if $\beta_s$ and $\beta_t$ are equal, then adding the source task dataset always helps learn the target task.
The goal of this section is to understand how covariate shift affects the rate of transfer. \todo{add conceptual msg}

%The key quantity is to look at:
%The estimator using the source and target together from minimizing \eqref{eq_mtl_basic} is
%\[ \hat{\beta}_{s,t} = (X_1^{\top} X_1 + X_2^{\top} X_2)^{-1} (X_1^{\top}Y_1 + X_2^{\top}Y_2)\]
%The estimation error of $\hat{\beta}_{s,t}$ is
%\begin{align}\label{eq_two_task}
%  \err(\hat{\beta}_{s,t}) = \sigma^2 \cdot \tr[(X_1^{\top}X_1 + X_2^{\top} X_2)^{-1}].
%\end{align}
%The estimation error using the target alone is
%\begin{align}\label{eq_target_task}
%	\err(\hat{\beta}_t) = \sigma^2 \cdot \tr[(X_2^{\top} X_2)^{-1}].
%\end{align}
%The improvement of estimation error from adding the source task is then given by
%$\err(\hat{\beta}_t) - \err(\hat{\beta}_{s,t})$.
%For the test error on the target task, the improvement from adding the source task is
%\[ \te(\hat{\beta}_t) - \te(\hat{\beta}_{s,t}) = \sigma^2\cdot\bigtr{\bigbrace{(X_2^{\top}X_2)^{-1} - (X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}\cdot\Sigma_2}. \]

%We calculate the amount of improvement by comparing equation \eqref{eq_two_task} to equation \eqref{eq_target_task}.
A simple observation here is that when $\beta_s = \beta_t$, the optimal $\hat{w}$ for minimizing equation \eqref{eq_te_mtl} is equal to $1$.
Based on this observation, we can get a more precise result than Theorem \ref{thm_model_shift} on the improvement of adding the source task data that only depends on the covariance matrices $\Sigma_1, \Sigma_2$ and the number of data points $n_1, n_2$.

\begin{proposition}[Transfer rate without model shift]\label{thm_cov_shift}
Suppose $\beta_s = \beta_t$ and $\|\beta_t\|_2^2\sim p\sigma^2$ (i.e. the $l^2$-norm of the vector $\beta_t$ is of the same order as that of the error vector). Assume that the condition numbers of $\Sigma_1$, $\Sigma_2$ and $M:=\Sigma_1^{1/2}\Sigma_2^{-1/2}$ are all bounded by a constant $C>0$. Then we have that the optimal ratio for $W_1/W_2$ in equation \eqref{eq_te_mtl} satisfies
	$$1\le \hat{w} \le 1+\OO(p^{-1}).$$%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
	%where $M:=\Sigma_1^{1/2}\Sigma_2^{-1/2}$.
Moreover, we have
	\begin{align}\label{tehatw1}
		%\err(\hat{\beta}^{\TL}_{s,t}) &= \sigma^2 \cdot \bigtr{\frac 1 {(n_1 + n_2)a_1\Sigma_1 + (n_1 + n_2)a_2\Sigma_2}} \\
		\te(\hat{\beta}^{\TL}_{t}) &= \sigma^2 \cdot \bigtr{\bigbrace{(n_1 + n_2)a_1 M^\top M  + (n_1 + n_2)a_2\id}^{-1}} \cdot \left(1+ \bigo{p^{-1}}\right),
	\end{align}
where $a_1, a_2$ are the solutions to equations \eqref{eq_a2}. %\cor $w_0$ is close to 1 if the signal strength $\beta_t$ is much larger than the noise strength \nc
\end{proposition}

\begin{proof}
We abbreviate $\val(w_2\hat{B}(w)):=\val(w)$. Note that $\val(w)\le \val(-w)$ for $w\ge 0$. Hence we have $\hat w\ge 0$. Moreover, we notice that $\val (w) < \val (1)$ for all $0\le w < 1$. Thus we have $\hat w\ge 1$. It suffices to consider the case with $w> 1$. Under the assumption on $\beta_s$ and $\beta_t$, we can write
\begin{align}
	\val(w) =&~  \left( 1-\frac1w\right)^2 \left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2 \nonumber \\
			&~ + \frac{\sigma^2}{w^2} \cdot \bigtr{( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} }. \nonumber
\end{align}
Since
\begin{align*}
&\left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2 \\
&= \tr \left[ ( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-2}M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t\beta_t^\top \Sigma_1^{1/2}Z_1^{\top} Z_1  M\right]
\end{align*}
is increasing with respect to $w$, then the derivative of $\val(w)$ can be bounded from below as
\begin{align*}
\val'(w) \ge &~ 2\frac{w-1}{w^3} \left\|( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2   \\
			&~ - 2 \frac{\sigma^2}{w^3} \cdot \bigtr{(M^\top Z_1^{\top}Z_1 M +w^{-2}  Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top}Z_1 M (M^\top Z_1^{\top}Z_1 M + w^{-2} Z_2^{\top}Z_2)^{-1}} \\
\ge &~ 2\frac{w-1}{w^3} \left\|( M^\top Z_1^{\top}Z_1 M +  Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2   - 2 \frac{\sigma^2}{w^3} \cdot \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}}.
			%\\ =& ~ 2\frac{d^2}{w^3} \tr\left[( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} \left[\left( w-1\right)\left(Z_1 \Sigma_1 Z_1^{\top}\right) - \frac{\sigma^2}{d^2}\id \right] Z_1 M ( M^\top Z_1^{\top}Z_1 M + w^{-2}Z_2^{\top}Z_2)^{-1}\right]  .
\end{align*}
Hence $\val'(w)\ge 0$ if $w>1+ \e_0$, where
$$\e_0:= \frac{\sigma^2  \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}} }{\left\|( M^\top Z_1^{\top}Z_1 M + Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2}.$$
In other words, $\val(w)$ is strictly increasing function on $[1+\e_0,\infty]$. Thus we get that $\hat w$ satisfies 
\be\label{hatw1}1\le w \le 1+\e_0.\ee Using \eqref{eq_isometric}, we get that
$$\e_0=\OO(\sigma^2/\|\beta_t\|_2^2)=\OO(p^{-1}).$$

Finally, plugging \eqref{hatw1} into the expression $\te(\hat{\beta}^{\TL}_{t}) $, we obtain \eqref{tehatw1}.
%$$1\le \hat{w} \le w_0:=1 +\frac{\sigma^2  \bigtr{(M^\top Z_1^{\top}Z_1 M)^{-1}} }{\left\|( M^\top Z_1^{\top}Z_1 M + Z_2^{\top}Z_2)^{-1} M^\top Z_1^{\top} Z_1 \Sigma_1^{1/2}\beta_t \right\|^2},$$
\end{proof}

As a remark, we see that Proposition \ref{prop_monotone} follows from Theorem \ref{thm_cov_shift}.
The amount of reduction on test error for the target task is given as
	\begin{align*}
%		\err(\hat{\beta}_t) - \err(\hat{\beta}_{s,t})
%		&= \sigma^2 p \cdot \bigtr{\frac 1 {(n_2 - p) \Sigma_2} - \frac 1 {(n_1 + n_2)a_1 \Sigma_1 + (n_1 + n_2)a_2 \Sigma_2}}, \\
		\te(\hat{\beta}_t) - \te(\hat{\beta}_{s,t})
		&= \sigma^2 \cdot \bigbrace{\frac p {n_2 - p} -  \bigtr{\bigbrace{(n_1 + n_2)a_1\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2} + (n_1 + n_2)a_2\id}^{-1}}}.
	\end{align*}
Because
\begin{align*}
	\te(\hat{\beta}_{s,t}) \le \te(\hat{\beta}_t)
	\Leftarrow~ & (n_2 - p)\Sigma_2 \preceq (n_1 + n_2) a_1 \Sigma_1 + (n_1 + n_2)a_2 \Sigma_2 \\
	\Leftrightarrow~ & \zeroMatrix \preceq (n_1 + n_2) a_1 \Sigma_1 + (n_1 - (n_1 + n_2)\cdot a_1) \Sigma_2,
\end{align*}
which is true since $a_1 \le n_1 / (n_1 + n_2)$ by equation \eqref{eq_a2}.
The proof for $\te(\hat{\beta}_{s,t}) \le \te(\hat{\beta}_t)$ follows by multiplying $\Sigma_2^{-1/2}$ on both sides of the inequalities above.

\medskip
Now we apply Theorem \ref{thm_cov_shift} to show how covariate shift affects the rate of transfer.

%\begin{example}[\textbf{When $\Sigma_1 = \Sigma_2$}]
%In this case, we have $\lambda_i = 1$ for all $1\le i\le p$.
%And $a_1 + a_2 = 1 - p / (n_1 + n_2)$.
%Hence
%\[ \te(\hat{\beta}_{s,t}) = \frac{\sigma^2 p^2}{n_1 + n_2 - p} \text{ and } \err(\hat{\beta}_{s,t}) = \frac {\sigma^2 p} {n_1 + n_2 - p} \bigtr{\Sigma_2^{-1}}. \]
%\end{example}

@@

\begin{figure}
	\centering
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/scaling.eps}
		\caption{When $\Sigma_1 = \Sigma_2 / \lambda$.}
		\label{fig_te_scaling}
	\end{minipage}\hfill
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/complementary.eps}
		\caption{When $\Sigma_1$ and $\Sigma_2$ are complementary. The number of target task data points is $n_2 = 4p$.}
		\label{fig_te_complement}
	\end{minipage}
\end{figure}

\paragraph{Extending the intuition to the general case.}
When there is model shift, i.e. $\beta_s = \beta_t$, we can still use Theorem \ref{thm_model_shift} (and Proposition \ref{prop_model_shift_tight}) to get the result.
\begin{itemize}
	\item \textbf{The effect of covariate shift:}
	\item \textbf{The effect of data ratio:}
\end{itemize}

>>>>

\subsection{Task Model Similarity versus Noise}

@@

\textbf{Implications.}

@@

Now we apply Theorem \ref{thm_cov_shift} to show how covariate shift affects the rate of transfer.

@@

\begin{figure}
	\begin{minipage}{0.48\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/complementary.eps}
		\caption{When $\Sigma_1$ and $\Sigma_2$ are complementary. The number of target task data points is $n_2 = 4p$.}
		\label{fig_te_complement}
	\end{minipage}
\end{figure}
